A brief follow-up from last time. We claimed that in $\text{AdS}_5\times S_5$, we had $T=0\implies R=0$ by the conformal symmetry. However, one might worry about the possibility of a trace anomaly in curved space. Normally, the trace of the stress tensor is proportional to $aE+\sum_i c_i C_i$, where $E$ is some Euler number term and the $C_i$s are constructed from many copies of the Riemannn tensor contracted with some $\epsilon$s. However, because $\text{AdS}_5\times S_5$ is conformally flat, all the $c_i$s vanish and moreover $E=0$ for odd-$D$ manifolds without boundary. Note also that $E(A\times B)=E(A)\times E(B)$. Hence $E=0$ for $\text{AdS}_5\times S_5$, so indeed there is no trace anomaly.

Note that the Newton constant of our spacetime depends on the compact Kaluza-Klein dimensions. Hence for instance (up to order unity factors)
\begin{equation}
    G_t \sim \frac{G_{10}}{\text{Area}[S_5]/l_{p,10}^5}.
\end{equation}
Hence gravity gets diluted in the presence of extra dimensions. In general $G\sim l_p^{D-2}$.

What evidence do we have that the AdS/CFT correspondence is true? In Maldacena's original derivation, he posited that the theory could be interpolated between weak and strong couplings so that the commuting square diagram really works.
\begin{itemize}
    \item The symmetries match up. Hence the $SO(2,4)\times SO(6)+32$ real supersymmetries line up. In $\cN=4$ SYM, we get Weyl spinors with 2 complex components. In the CFT, there are 16 supercharges $Q_A^\alpha$ and 16 superconformal $S_{\bar A}^\alpha = X_{A\bar A} Q_B^\alpha \epsilon^{AB}$.%superconformal what? charges?
    There are some anticommutation relations which allow us to construct the representations of the symmetries,
    \begin{align*}
        \set{Q_A,\bar Q_{\bar A}} &\sim P_{A\bar A}\\
        \set{S,\bar S} &\sim K\\
        \set{Q,\bar S} &\sim M+R+D,
    \end{align*}
    where $P$ are the $+1$ spin representations, $Q$ the $+1/2$, $R,M,D$ the $0$ reps, $S$ the $-1/2$ reps, and $K$ the $-1$ reps.
    \item S-duality. $SL(2,\ZZ)$: in SYM, we have $\tilde g=g + i\theta, \theta F \wedge F$ and on SUGRA, $g_5:\Phi,C_0$
    \item Trace anomalies match (protected from weak$\to$strong coupling)
    \item BH entropy/holographic entanglement entropy (to be discussed more alter)
    \item ``chiral'' primaries of supergroup-- irrep that has null states, $Q\ket{\psi}=0$ for some $Q,\psi$, with $S\ket{\psi}=0$ and $\set{Q,S}\ket{\psi}=0$. $\Delta$ is determined by spin ($M$) and $R$-charge ($R$), hence the charge $Q$ is protected in going to strong coupling.
\end{itemize}

In fact, all type IIB supermultiplets (KKK reduced to AdS${}_5$ in regime of validity of SUGRA) is equivalent to the single trace chiral irreps which are built $\text{tr}(\underbrace{\phi_{\{I}\phi_J \phi_{K\ldots\}}}_{n})$ where $\set{}$ indicates symmetric traceless. This single trace expression corresponds to the $n$th spherical harmonic on $S_5$. We then have the relations
\begin{align*}
    [Q_A,\phi]&\sim \psi_A,\\
    \set{Q_A,\psi_B} &\sim F_{AB} + [\phi^I,\phi^J]\epsilon_{AB}\\
    \set{Q_A,\psi_{\bar B}^*} & \sim D_{A\bar A} \phi\\
    [Q_A,A_{B\bar B}]&\sim \epsilon_{AB}\psi_{\bar B}^*.
\end{align*}

We can have traces of $2\leq n \leq N$ $\phi$s-- note that traces of more than $N$ commuting $N\times N$ matrices are not independent (of traces of fewer than $N$ matrices). This relation suggests that there is a shortest scale on the $S_5$, such that no new degrees of freedom emerge if we try to probe length scales smaller than the Planck scale.

Now, IIB has no additional fields, so it seems that the non-chiral irreps have $\Delta \to$large ($\Delta^2 \sim 1/l_s$) as $\lambda\to$large. In $n=2$ we have $\text{tr}(\phi_{\{I}\phi_{J\}})$ with $2\leq \Delta \leq 4$ which includes all conserved CFT currents. $\Delta=3=d-1$ is the $R$-charge $J$, $\Delta=3+1/2$ gives $J_\text{super}$, and $\Delta=4=d$ gives $\Phi$ dual to the dilaton. %??

There are other versions of the duality which are useful.
\begin{itemize}
    \item $N$ M2's: $\text{AdS}_4\times S_6^{(m)}\longleftrightarrow d=5, \cN=8$ ABJM model. $U(N)\times U(N)$ Chern-Simons gauge theory $\cL= F\wedge A +\frac{2}{3} A \wedge A \wedge A$, also IR limit of $d=3, \cN = 8$ SYM ($g=$relevant) (D2)
    \item $N$ M5s: $\text{AdS}_2 \times S_4 \longleftrightarrow d=6, \cN=(2,0)$ max dimensional superconformal theory. Does not come from a Lagrangian but is known to exist from stringy arguments. Compactify on $S_1$, flows to $d=5,\cN=4$ SYM ($g=$irrelevant) (D4)
    \item $Q_1$ D1 $Q_5$: $\text{AdS}_3 \times S_3 \times T_4$ (IIB) (K3) $\longleftrightarrow$ nonlinear $\sigma$-model $d=2$.
\end{itemize}