A quick bit of admin-- office hours are rescheduled to tomorrow (February 13) from 2-4 PM.

Last time, we wrote down a 1-parameter family of Hamiltonians
\begin{equation*}
    \hat H =-\frac{1}{2}\paren{\frac{1}{\sqrt{g}} \P{}{x^a} \paren{g^{ab}\sqrt{g}\P{}{x^b}} + \alpha R[g]},
\end{equation*}
where this first term is none other than $-\frac{1}{2}\nabla^a \nabla_a$, a covariant Laplacian. However, we also get this extra term-- we get a bit of coupling to the Ricci scalar which we're free to choose.

\subsection*{Supersymmetric NLSM}
As always, our goal will be to supersymmetrize the model and see what we learn. Take a bosonic variable $x$ and fermionic variable $\psi^a$ such that
\begin{equation}
    x: M\to N, \quad \psi^a \in \prod \Omega^0(M,x^* T_N)
\end{equation}
where $N$ is some target Riemannian manifold $(N,g)$, $M$ is something like an interval $[0,\beta]$ or $S^1$, and $\Omega^0$ is a function on the worldline.

With these variables, we can write down an action
\begin{equation}
    S[x,\psi]=\int_M \bkt{\frac{1}{2} g_{ab} \dot x^a \dot x^b + i g_{ab} \bar \psi^a(\nabla_t \psi)^b -\frac{1}{2} R_{abcd} \psi^a \bar \psi^b \psi^c \bar \psi^c} dt
\end{equation}
where $\nabla_t\psi^a = \frac{d\psi^a}{dt}+\Gamma^a_{bc} \frac{dx^b}{dt} \psi^c$ is the pullback of the connection on $N$. This first term is just our bosonic kinetic term, and the second is a reasonable sort of fermionic kinetic term, but the last is a bit of a surprise. The fermions couple directly to the Riemann curvature of the target space $N$.

Note that this action is invariant under the SUSY transformations
\begin{gather}
    \delta x^a = \epsilon \bar \psi^a -\bar \epsilon \psi^a\\
    \delta \psi^a = \epsilon(i\dot x^a - \Gamma^a_{bc} \bar \psi^b \psi^c)\\
    \delta \bar \psi^a = \bar \epsilon(-i\dot x^a - \Gamma^a_{bc} \bar \psi^b \psi^c),
\end{gather}
which we can check explicitly (e.g. on the second examples sheet). As before, the origin of these transformations may seem a bit mysterious but will become clearer when we discuss superfields.

These transformations are generated by the Noether charges $Q,\bar Q$, where
\begin{align}
    Q &= i\bar \psi^a(g_{ab} \dot x^b +ig_{bc} \bar \psi^b \Gamma^c_{ad} \psi^d),\\
    \bar Q &= -i\psi^a(g_{ab} \dot x^b + i g_{bc} \psi^b \Gamma^c_{ad} \bar \psi^d).
\end{align}

Our action is also invariant under the (considerably simpler) transformation $\psi^a \mapsto e^{i\alpha}\psi^a, \bar \psi^a \mapsto e^{-i\alpha}\bar \psi^a$, generated by the charge $F=g_{ab} \psi^a \bar \psi^b$. Conservation of $F$ in the quantum theory tells us that no fermionic excitations are created or destroyed by time evolution.

Quantizing this theory, we have the conjugate momenta
\begin{equation}
    p_a= \frac{\delta L}{\delta \dot x^a}=g_{ab}\dot x^b + i g_{bc} \bar \psi^b \Gamma^c_{ad} \psi^d; \quad \pi_a =\frac{\delta L}{\delta \dot \psi^a}=ig_{ab} \bar \psi^b.
\end{equation}
These have canonical (anti)commutation relations
\begin{equation}
    [\hat x^a,\hat p_b]=i\delta^a{}_b,\quad \set{\hat \psi^a,\hat{\bar\psi}^b}=g^{ab}(x),
\end{equation}
with all others trivial. For the bosonic fields, we choose the Hilbert space
\begin{equation*}
    \cH =L^2(N,\sqrt{g}d^nx),
\end{equation*}
with $\hat p_a \to -i\P{}{x^a}$. For the fermions, we again choose $\bar \psi^a$ to be raising operators and $\psi^a$ lowering operators. If we then pick a vacuum $\ket{0}$ defined by $\psi^a\ket{0}=0 \forall a$, all other states of the fermionic system are generated by acting with the $\bar \psi$s on $\ket{0}$.

Notice that each $\bar \psi^a$ can only act once, since $\set{\bar \psi^a, \bar \psi^b}=0$. This is because our whole universe is just an interval (remember, $M=[0,\beta]$ or $S^1$), so rather than having Fourier modes defined on a time slice, we have different fields defined at points. This enforces the Pauli exclusion principle, if you like. However, we do get different fermions since we have an index $\bar \psi^a$ to range over. We can interpret these as \emph{forms} on $N$: thus
\begin{align}
    \ket{0} &\leftrightarrow 1\\
    \bar \psi^a \ket{0} &\leftrightarrow dx^a\\
    \bar \psi^a \bar \psi^b \ket{0} &\leftrightarrow dx^a \wedge dx^b\\
    \bar \psi^a \ldots \bar \psi^n \ket{0} &\leftrightarrow dx^1 \wedge dx^2 \wedge \ldots \wedge dx^n.
\end{align}
There are no more since any $(\bar \psi^a)^2=0$.

Altogether, the Hilbert space of SUSY QM is thenrefore
\begin{equation}
    \cH = \Omega^\cdot (N)=\bigoplus_{p=0}^n \Omega^p(N)
\end{equation}
where $\Omega^p(N)$ is the space of $p$-forms on $N$, i.e. a general state $\Psi(x,\bar \psi)$ can be written as
\begin{align}
    \Psi(x,\bar \psi)&=f(x) + \alpha_a(x) \bar \psi^a + \beta_{ab}(x) \bar \psi^a \bar \psi^b +\ldots + \omega_{1\ldots n}(x) \bar \psi^1 \ldots \bar \psi^n\\
    &\sim f(x) + \alpha_a dx^a + \beta_{ab} dx^a \wedge dx^b + \ldots + \omega dx^a \wedge \ldots \wedge dx^n.
\end{align}
So there is a direct correspondence between expanding in $\bar \psi$ operators and in terms of differential forms.

Acting on this space, we can set up the full correspondence
\begin{gather*}
    \hat x^a \to x^a \times(\cdot)\\
    \hat p_a \to -i \P{}{x^a}(\cdot)\\
    \bar \psi^a \to dx^a \wedge(\cdot)\\
    \psi^a \to g^{ab}\P{}{x^b}(\cdot)\text{ (contraction)}.
\end{gather*}
Thus when we hit some arbitrary state $\bar \psi^a \bar \psi^b \bar \psi^c \ldots \bar \psi^d\ket{0}$ with a lowering operator (where there are an odd number of $\bar \psi$s), we can use the anticommutation relations to find
\begin{align*}
    \psi^e(\bar \psi^a \bar \psi^b \bar \psi^c \ldots \bar \psi^d \ket{0}) &= \set{\psi^e, \bar \psi^a \bar \psi^b \ldots \bar \psi^d}\ket{0}\\
        &= \paren{
            \set{\psi^e,\bar \psi^a}\bar \psi^b \ldots \bar \psi^d - \bar \psi^a \set{\psi^e,\bar \psi^b}\bar \psi^c \ldots \bar \psi^d + \ldots +\bar \psi^a \bar \psi^b \ldots \set{\psi^e,\bar \psi^d}
        }\ket{0}\\
        &= \paren{g^{ea}\bar \psi^b \ldots \bar \psi^d - g^{eb} \bar \psi^a \bar \psi^c \ldots \bar \psi^d + \ldots + g^{ed} \bar \psi^a \bar \psi^b \ldots} \ket{0}.
\end{align*}
But notice this is just what we would have gotten from contracting
\begin{equation}
    {}^\iota g^{ef} \P{}{x^f}(dx^a \wedge dx^b \ldots \wedge dx^d).
\end{equation}
The inner product on $\cH$ is 
\begin{equation}
    \braket{\alpha}{\beta} =\int_N \alpha \wedge * \beta,
\end{equation}
where $*$ is the Hodge star operator.%
    \footnote{No, we didn't learn enough of this differential geometry in \emph{General Relativity}. Blame Malcolm Perry. I direct you to Harvey Reall's GR notes and/or Sean Carroll's excellent textbook \emph{Spacetime and Geometry} as a reference.}
Here, if $\alpha,\beta\in \Omega^p(N)$ ($p$-forms on $N$), then
\begin{equation}
    \int_N \alpha \wedge * \beta = \int_N \alpha^{a_1\ldots a_p}\beta_{a_1\ldots a_p} \sqrt{g},
\end{equation}
with indices raised using $g^{ab}$. For $\alpha \in \Omega^p(N), \beta \not\in \Omega^p(N),$ we simply define $\int_N \alpha \wedge * \beta =0$. This follows since $\psi^a$ is the adjoint of $\bar \psi^a$, so
\begin{align}
    \int_N \overline{\alpha}_{a_1 \ldots a_p}(x)\beta_{b_1\ldots b_p}(x) \sqrt{g} d^n x &{} \underbrace{\bra{0} \psi^{a_1}\ldots \psi^{a_p}}_{\text{fermionic part of }\bra{\alpha}} \underbrace{\bar \psi^{b_1} \ldots \bar \psi^{b_p} \ket{0}}_{\text{fermionic part of }\ket{\beta}}
    &=\int_N \overline{\alpha}^{a_1\ldots a_p}\beta_{a_1\ldots a_p}\sqrt{g} d^n x.
\end{align}
We see that the integral vanishes unless we precisely match the indices between $\alpha$ and $\beta$.

Furthermore, in the quantum theory we get
\begin{equation}
    Q=i\bar \psi^a \hat p_a \to dx^a \P{}{x^a} = d,
\end{equation}
where $d$ is now the exterior derivative, taking us from $\Omega^p(N)\to \Omega^{p+1}(N)$, from $p$-forms to $p+1$-forms.
Similarly,
\begin{equation}
    \bar Q = -i\psi^a \hat p_a \to d^\dagger,
\end{equation}
where $d^\dagger$ is the adjoint with respect to the inner product $\avg{{},{}}$ which takes us from $\Omega^p(N)\to \Omega^{p-1}(N)$, from $p$ to $p-1$-forms, such that if $\alpha\in \Omega^p, \beta\in \Omega^{p+1},$ then $\avg{\alpha,d^\dagger\beta}=\avg{d \alpha, \beta}.$