Today we'll finish our discussion of the zero-dimensional path integral by introducing fermions to our theory. To model fermions, we will introduce Grassmann variables,%
    \footnote{We've seen these in \emph{Supersymmetry} already.
    }
i.e. a set of $n$ elements $\set{\theta_a}_{a=1}^n$ obeying anticommutation relations,
\begin{equation}
    \theta_a \theta_b = -\theta_b \theta_a.
\end{equation}
Note also that for (complex) scalars $\phi_b\in \CC$,
\begin{equation}
    \theta_a \phi_b = \phi_b \theta_a,
\end{equation}
i.e. scalars commute with Grassmann variables. In addition, $\theta^2_a =0$ by the anticommutation relations, which implies that any function of $n$ Grassmann variables can be written in finite form. That is, polynomials in Grassmann variables are forced to terminate since at some point we run out of distinct Grassmann variables to multiply. A general function $F(\theta)$ can be written
\begin{equation}
    F(\theta)=f+\rho_a \theta_a +\frac{1}{2!} g_{ab} \theta_a \theta_b + \ldots + \frac{1}{n!} h_{a_1\ldots a_n} \theta_{a_1}\ldots \theta_{a_n}.
\end{equation}
Note that the coefficients $\rho,g,\ldots,h$ are totally antisymmetric under interchange of indices.

We also want to define differentiation and integration of these guys. Differentiation anticommutes with the Grassmann variables, i.e.
\begin{equation}
    \paren{\P{}{\theta_a}\theta_b + \theta_b \P{}{\theta_a}} * = (\delta_{ab})*
\end{equation}
where the derivative in the first term acts on everything coming after. This leads us to a modified Leibniz rule.

To define integration, note that for a single Grassmann variable $\theta$, a function takes the form
\begin{equation}
    F(\theta)=f+\rho \theta,
\end{equation}
so we just need to define $\int d\theta$ and $\int d\theta \,\theta$. If we require translational invariance, i.e.
\begin{equation}
    \int d\theta(\theta+\eta)=\int d\theta \theta \implies \int d\theta =0.
\end{equation}
We can then choose the normalization so that $\int d\theta \, \theta = 1$. Note the similarity between differentiation and integration (i.e. an integral $\int d\theta\,\theta =\P{}{\theta}\theta=1$). This process is called \term{Berezin integration}. Using these rules, we also find that
\begin{equation}
    \int d\theta \P{}{\theta} F(\theta)=0,
\end{equation}
since the term linear in $\theta$ will go to a constant by the derivative and be killed by the integral, and any constant terms will be killed by the derivative. Either way the result is zero.

Suppose now we have $n$ Grassmann variables. Then the only nonvanishing integrals involve exactly one power of each integration variable, e.g.
\begin{equation}
    \int d^n \theta\, \theta_1 \theta_2 \ldots \theta_n = \int d\theta_n d\theta_{n-1}\ldots d\theta_1 \, \theta_1 \theta_2 \ldots \theta_n = 1.
\end{equation}
In general we can just anticommute the Grassmann variables until they're in the right order, picking up a factor for the parity of the permutation. That is,
\begin{equation}
    \int d^n\theta \theta_{a_1}\theta_{a_2}\ldots \theta_{a_n} = \epsilon^{a_1 a_2 \ldots a_n},
\end{equation}
where $\epsilon$ is the totally antisymmetric symbol with value $+1$ for even permutations of $1,2,\ldots,n$, $-1$ for odd permutations, and $0$ if any indices are repeated.

What if we now make a change of variables $\theta_a' = A_{ab} \theta_b$? Then
\begin{align}
    \int d^n \theta \theta'_{a_1} \theta'_{a_2} \ldots \theta'_{a_n} &= A_{a_1b_1}\ldots A_{a_nb_n} \underbrace{\int d^n \theta \, \theta_{b_1} \ldots \theta_{b_n}}_{\epsilon^{b_1\ldots b_n}}\\
    &= \det A \,\epsilon^{a_1\ldots a_n}\\
    &= \det A \int d^n \theta' \,\theta'_{a_1} \ldots \theta'_{a_n}
\end{align}
We conclude that under a change of variables, the integration measures are related by
\begin{equation}
    d^n\theta = \det A \,d^n \theta'.
\end{equation}
Note that this is the opposite of the convention for scalars, where
\begin{equation}
    \phi'_a = A_{ab} \phi_b \implies d^n \phi =\frac{1}{|\det A|}d^n \phi'.
\end{equation}

\subsection*{Free fermion field theory} Consider $d=0$, with two fermion fields $\theta_1,\theta_2$. The action must be bosonic (scalar), so the only possible nonconstant action is
\begin{equation}
    S(\theta)=\frac{1}{2}A \theta_1 \theta_2, A\in \RR
\end{equation}
Then the path integral is
\begin{equation}
    Z_0 = \int d^2 \theta \, e^{-S(\theta)/\hbar}=\int d^2\theta \paren{ 1-\frac{A}{2\hbar}\theta_1\theta_2} = -\frac{A}{2\hbar},
\end{equation}
where the exponential has terminated thanks to our Grassmann variables.

Suppose now we have $n=2m$ fermion fields $\theta_a$. Then our action might be quadratic in the fields,
\begin{equation}
    S=\frac{1}{2} A_{ab} \theta_a \theta_b
\end{equation}
with $A$ an antisymmetric matrix, and the path integral is then
\begin{align*}
    Z_0 &= \int d^{2m}\theta\, e^{-S(\theta)/\hbar} = \int d^{2m} \theta \sum_{j=0}^{m} \frac{(-1)^j}{(2\hbar)^j j!} (A_{ab}\theta_a \theta_b)^j\\
    &= \frac{(-1)^m}{(2\hbar)^m m!} \int d^{2m}\theta A_{a_1 a_2} A_{a_3 a_4} \ldots A_{a_{2m-1} a_{2m}} \theta_{a_1} \theta_{a_2} \ldots \theta_{2m}\\
    &= \frac{(-1)^m}{(2\hbar)^m m!} \epsilon^{a_1 a_2 \ldots a_{2m}} A_{a_1 a_2} A_{a_3 a_4} \ldots A_{a_{2m-1} a_{2m}}\\
    &= \frac{(-1)^m}{\hbar^m} \text{Pf}(A),
\end{align*}
where $\text{Pf}(A)$ is the \term{Pfaffian} of the matrix $A$, defined by
\begin{equation}
    \text{Pf}(A)\equiv \frac{1}{2^m} \epsilon^{a_1 a_2 \ldots a_{2m}} A_{a_1 a_2} A_{a_3 a_4} \ldots A_{a_{2m-1} a_{2m}},
\end{equation}
which we will show on the examples sheet is in fact $\pm \sqrt{\det A}.$ Thus $\text{Pf}\begin{pmatrix}0 & -1 \\ a & 0 \end{pmatrix} = a$. Using this property, we find that for fermionic fields,
\begin{equation}
    Z_0 = \pm \sqrt{\frac{\det A}{\hbar^n}}
\end{equation}
with $A$ antisymmetric, whereas for bosonic fields with some symmetric mass matrix $M$,%
    \footnote{That is, for an action $S=\frac{1}{2}M_{ab}\phi_a \phi_b$.}
we have
\begin{equation}
    Z_0=\sqrt{\frac{(2\pi \hbar)^n}{\det M}}.
\end{equation}

We can now introduce an external source function to our action, a Grassmann-values $\set{\eta_a}$, such that the new action is
\begin{equation}
    S(\theta,\eta)=\frac{1}{2} A_{ab} \theta_a \theta_b + \eta_a \theta_b.
\end{equation}
Taking care to respect the anticommutation relations and completing the square as before, we can rewrite the action as
\begin{equation}
    S(\theta,\eta)=\frac{1}{2}(\theta_a +\eta_c(A^{-1})_{ca}) A_{ab}(\theta_b +\theta_d(A^{-1})_{db}) +\frac{1}{2} \eta_a (A^{-1})_{ab} \eta_b.
\end{equation}
We can make a change of variables using the translational invariance of $\theta_a$ and pull out the constant factor to find
\begin{equation}
    Z_0(\eta)=\exp\paren{-\frac{1}{2\hbar}\eta^T( A^{-1}) \eta} Z_0(0).
\end{equation}
This allows us to get propagators by taking derivatives with respect to the source $\eta$, as we are wont to do:
\begin{equation}
    \avg{\theta_a \theta_b}
    = \frac{\hbar^2}{Z_0(0)}\frac{\p^2 Z_0(\eta)}{\p \eta_a \p \eta_b}|_{\eta=0} 
    = \hbar(A^{-1})_{ab}.
\end{equation}
We see that the propagator is proportional to the inverse of the bilinear part of the action for Grassmann variables.