Last time, we finished discussing the representation theory of $L(SU(2))$. In particular, we defined the tensor product representation and showed that we can usually express the tensor product of two representations in terms of the direct product of many copies of $R_{\Lambda}$:
$$R_\Lambda \otimes R_{\Lambda'}=\bigoplus_{\Lambda'' \in \ZZ_{\geq 0}} L^{\Lambda''}_{\Lambda,\Lambda'}R_{\Lambda''}$$
with $$L^{\Lambda''}_{\Lambda,\Lambda'}\in \ZZ_{\geq 0}.$$

We also described an algorithm to work out the direct product representation, namely writing the tensor product as a direct sum of the representation $R_{\Lambda+\Lambda'}$ and some remainder term $\tilde R_{\lambda,\Lambda'}.$ It's an exercise (sheet 2, Q3) to work out that
$$R_N\otimes R_M = R_{|N-M|}\oplus R_{|N-M|+2} \oplus\ldots \oplus R_{N+M}.$$
Tensor products are important because multi-particle spaces are described in general by tensor products, not direct products (this leads to the phenomenon of entanglement).

Let us now define something called the \emph{Killing form.}
\begin{defn}
Given a vector space $V$ over $F$ ($=\RR,\CC$) an \term{inner product} $i$ is a symmetric bilinear map
$$i:V\times V \to F.$$
In particular, $i$ is \term{non-degenerate} if for every $v\in V$ ($v\neq 0$), there is a $w\in V$ such that
$$i(v,w)\neq 0.$$ That is, there is no vector that is orthogonal to all the others under the inner product, or equivalently it has no zero eigenvalues considered as a linear map.
\end{defn}

Question: is there a ``natural'' inner product on a Lie algebra $\fg$? The answer is yes-- it is called the \term{Killing form}, an inner product $\kappa$ with
$$\kappa:\fg \times \fg \to F.$$
We'll define the formula first and then explore why it makes sense.
\begin{defn}
The Killing form $K$ is defined such that $\forall X,Y \in \fg$,
$$\kappa(X,Y)\equiv \text{Tr}(\text{ad}_X \circ \text{ad}_Y).$$
That is, $K$ is the trace of the linear map
$$\text{ad}_X \circ \text{ad}_Y: \fg \to \fg$$
which takes
$$Z\in \fg \mapsto [X,[Y,Z]] \in \fg.$$
\end{defn}
Why is this a sensible choice? Suppose we choose a basis $\set{T^a},a=1,\ldots D$ for $\fg$ with dimension $D$. Then
$$X=X_a T^a,\quad Y=Y_a T^a, \quad Z=Z_a T^a.$$
We also have some structure constants associated to the basis,
$$[T^a,T^b]=f^{ab}_c T^c.$$
Thus the composition of the ad maps is some $D\times D$ matrix, and we can work out in this basis the components of this matrix.
\begin{align*}
[X,[Y,Z]]&= X_a Y_b Z_c [T^a,[T^b,T^c]]\\
&= X_a Y_b Z_c [T^a,f^{bc}_d T^d]\\
&= X_a Y_b Z_c f^{ad}_e f^{bc}_d T^e\\
&= M(X,Y)^c_e Z_c T^e
\end{align*}
with 
$$M(X,Y)^c_e \equiv X_a Y_b f^{ad}_c f^{bc}_d.$$
The matrix $M(X,Y)$ is therefore the linear map $\text{ad}_X\circ \text{ad}_Y:\fg \to \fg$, and all that remains is to take the trace to get the Killing form.
\begin{align*}
\kappa^{ab}X_a Y_b &= \text{Tr}_D[M(X,Y)]\\
&= M(X,Y)^c_c\\
&= X_a Y_b f^{ad}_c f^{bc}_d.
\end{align*}
Therefore the Killing form in terms of structure constants is explicitly
$$\kappa^{ab}=f^{ad}_c f^{bc}_d.$$
The indices $c$ and $d$ are summed over, so we get the two free indices $a,b$ as desired.

Now what do we mean by saying that the Killing form is a ``natural'' inner product on a Lie algebra? It is the property that $\kappa$ is invariant under the adjoint action of $\fg$,
$$\kappa([Z,X],Y)+\kappa(X,[Z,Y])=0$$ for all $Z\in \fg, X,Y \in \fg.$ This is the equivalent of invariance under a conjugation by a Lie algebra element, $g X g^{-1}.$

Let's show that this property holds for this inner product.
\begin{align*}
\kappa([Z,X],Y) &=\text{Tr}[\text{ad}_{[Z,X]}\circ \text{ad}_Y]\\
&=\text{Tr}[(\text{ad}_Z \circ \text{ad}_X - \text{ad}_X \circ \text{ad}_Z)\circ \text{ad}_Y]\\
&=\text{Tr}[\text{ad}_Z \circ \text{ad}_X \circ \text{ad}_Y] - \text{Tr}[\text{ad}_X \circ \text{ad}_Z \circ \text{ad}_Y]
\end{align*}
where in going from the first to the second line, we have used the fact that the ad map is also a representation and can therefore be rewritten by linearity in its argument $[Z,X]$.
Similarly,
$$\kappa(X,[Z,Y])=\text{Tr}[\text{ad}_X\circ \text{ad}_Z \circ \text{ad}_Y]-\text{Tr}[\text{ad}_X \circ \text{ad}_Y \circ \text{ad}_Z].$$
However, if we now compare these two expressions we see that by the cyclic property of the trace (i.e. interpreting the ad maps as matrices on the vector space), their sum vanishes\footnote{Explicitly, this means that $\text{Tr}[\text{ad}_X \circ \text{ad}_Y \circ \text{ad}_Z]=\text{Tr}[\text{ad}_Z \circ \text{ad}_X \circ \text{ad}_Y]$.}, and so
$$\kappa([Z,X],Y)+\kappa(X,[Z,Y])=0. \qed$$

We may next ask under what conditions $\kappa$ is non-degenerate, i.e. the map $\kappa^{ab}$ is invertible. 
\begin{thm}
(Cartan) The Killing form $\kappa$ on a Lie algebra $\fg$ is non-degenerate $\iff \fg$ is semi-simple.
\end{thm}
In the specific case, if $\fg$ is simple, then the Killing form $\kappa$ is the unique invariant inner product on $\fg$ up to an overall scalar multiple.
\begin{defn}
A Lie algebra is \term{semi-simple} if it has no abelian ideals. (This is a little weaker than simple, clearly.)
\end{defn}
\begin{ex}
From Example Sheet 2, Question 9b: Show that a finite dimensional semi-simple Lie algebra can be written as the direct sum of a finite number of simple Lie algebras,
$$\fg = g_1\oplus g_2 \oplus \ldots \oplus g_l, \quad g_i\text{ simple}.$$
Note that a direct product $\fg \oplus \mathfrak{f}$ of Lie algebras $\fg,\mathfrak{f}$ is defined such that $\forall X\in \fg, Y \in \mathfrak{f}, [X,Y]=0.$
\end{ex}

Let us prove the forward direction of Cartan's theorem. First note that $$\kappa\text{ non-degenerate }\implies \fg\text{ is semi-simple}$$ is equivalent to proving the contrapositive, 
$$\fg\text{ not semi-simple }\implies \kappa\text{ is degenerate}.$$
Suppose $\fg$ is not semi-simple. Then $\fg$ has an abelian ideal $\mathfrak{j}.$ Let $\dim(\fg)=D$ and suppose the ideal has dimension $\dim(\mathfrak{j})=d.$ WLOG we can choose a basis $B$ for $\fg$ such that
$$B=\set{T^a} =\set{\underbrace{T^i; i=1,\ldots,d}_{\text{span }\mathfrak{j}}}\cup \set{T^\alpha;\alpha=d+1,d+2,\ldots, D},$$
i.e. a subset $T^i$ of the basis vectors span the ideal $\mathfrak{j}.$ Since $\mathfrak{j}$ is abelian,
$$[T^i,T^j]=0\quad\forall i,j=1,\ldots,d.$$ Therefore the structure constants are constrained by
$$f^{ij}_a = 0,\quad i,j=1,\ldots, d, a=1,\ldots,D.$$
That is, the bracket vanishes for all pairs of elements in the abelian ideal, so all structure constants with $ij$ indices up are zero.

Moreover, $\mathfrak{j}$ is an ideal, so the bracket of a basis element for $\mathfrak{j}$ with a general basis element is still in $\mathfrak{j}.$ That is,
$$[T^\alpha,T^j]=f^{\alpha j}_k T^k \in j \implies f^{\alpha j}_\beta=0, \beta = d+1,\ldots,D.$$
We'll use these facts next time to complete the proof of Cartan's theorem in one direction.