Admin note: there was no lecture (and hence no notes) for Friday, February 8, as Prof. Datta sustained an injury which prevented her from giving the lecture.

\subsection*{Quantum operations and CPTP maps}
To recap from last time, any allowed physical process on a quantum system is given by a quantum operation. The map must be completely positive (CP) in order to allow us to properly couple an ancilla (environment) to our system, and it must be linear and trace-preserving in order to take density matrices to other density matrices.

Consider a map $\Lambda:B(\cH)\to B(\mathcal{K}),$ where $\cH\simeq \CC^m,\mathcal{K}\simeq \CC^n$. Let $\cM_m$ be the space of $m\times m$ complex semi-definite matrices, and $\cM_m^+$ the same but restricted to positive matrices. The set of density matrices on $\CC^n$ is given by
\begin{equation}
    \cD(\CC^m)=\set{\rho \in \cM_m^+; \Tr \rho = 1}.
\end{equation}

\begin{defn}
    A map $\Lambda:\cM_m\to \cM_n$ is positive if
    \begin{equation}
        \Lambda(A)\in \cM_n^+ \forall A \in \cM_m^+.
    \end{equation}
\end{defn}
\begin{defn}
    For a given positive integer $k$, $\Lambda$ is $k$-positive if
    $(\Lambda \otimes \text{id}_k)$ is positive, where $\text{id}_k$ is the identity (super)operator, $\text{id}_k:\cM_k \to \cM_k$ such that $\text{id}_k(Q)=Q \forall Q\in \cM_k$.
\end{defn}
\begin{defn}
    The map $\Lambda$ is \term{completely positive} (CP) if it is $k$-positive $\forall k\in \ZZ^+$.
\end{defn}
\begin{thm}[Necessary and sufficient condition for CP]
    A linear map $\Lambda:\cB(\CC^d)\to \cB(\CC^{d'})$ is completely positive $\iff (\Lambda \otimes \id_d)(\kb{\Omega}{\Omega})\geq 0$, where
    \begin{equation}
        \ket{\Omega}=\frac{1}{\sqrt{d}}\sum_{i=1}^d \ket{i}\ket{i} \in \CC^d \otimes \CC^d.
    \end{equation}
\end{thm}
That is, it suffices to check positivity on the density matrix corresponding to the maximally entangled $d$-dimensional state.
\begin{proof}
    Necessity follows immediately from the definition of CP. To show sufficiency, consider an arbitrary $k\geq 1$. For a state $\rho \in \cD(\CC^d \otimes \CC^k)$, we have a spectral decomposition
    \begin{equation}
        \rho =\sum \lambda_i \kb{\phi_i}{\phi_i}
    \end{equation}
    where $\ket{\phi_i}\in \CC^d \otimes \CC^k.$
    Now we have
    \begin{align}
        (\Lambda \otimes \text{id}_k)\rho \geq 0 &\implies \sum_i \lambda_i(\Lambda \otimes \id_k)(\kb{\phi_i}{\phi_i})\geq 0\\
        &\implies \forall i, (\Lambda \otimes \id_k)\kb{\phi_i}{\phi_i} \geq 0.
    \end{align}
    We saw that for each of the basis states $\ket{\phi_i}$, we could write it as
    \begin{equation}
        \ket{\phi_i}=(I\otimes R_i)\ket{\Omega}
    \end{equation}
    for some $R_i\in \cB(\CC^d, \CC^k)$. Thus we can rewrite the basis states in our inequality to get
    \begin{equation}
        (\Lambda \otimes \id_k)(I\otimes R_i)\kb{\Omega}{\Omega} (I\otimes R_i^\dagger)\geq 0.
    \end{equation}
    Note that with the following definition
    \begin{equation}
        (\id_d \otimes f_i)(\omega) := (I\otimes R_i)(\omega(I\otimes R_i^\dagger)),
    \end{equation}
    our inequality becomes
    \begin{equation}
        (\Lambda \otimes \id_k)(\id_d \otimes f_i)(\dyad{\Omega}) \geq 0.
    \end{equation}
    Rewriting, this expression becomes
    \begin{equation}
        (\id_{d'}\otimes f_i)(\Lambda \otimes \id_d)\kb{\Omega}{\Omega} = \underbrace{(I_{d'}\otimes R_i)}_A\underbrace{(\Lambda \otimes \id_d)(\kb{\Omega}{\Omega})}_{B}\underbrace{(I_{d'}\otimes R_i^\dagger)}_{A^\dagger}.
    \end{equation}
    This is equivalent to the condition on matrices that $ABA^\dagger \geq 0$, and it turns out that for $ABA^\dagger \geq 0$, it suffices to have $B\geq 0$.%
        \footnote{Basically, if $B\geq 0$ then $\bra{v}B \ket{v}\geq 0\forall v$. But then define $A^\dagger v' = v,$ and we see that 
        \begin{equation*}
            \bra{v}B\ket{v}=\bra{A^\dagger v'}B\ket{A^\dagger v'}=\bra{v'}A BA^\dagger \ket{v'}\geq 0 \,\forall v'.
        \end{equation*}
        }
    Thus
    \begin{equation}
        (\Lambda \otimes \id_d)\kb{\Omega}{\Omega} \geq 0. \qedhere
    \end{equation}
\end{proof}

This construction we have defined is known as the \term{Choi matrix} (a Choi state of $\Omega$), i.e.
\begin{equation}
    J\equiv J(\Lambda) = (\Lambda \otimes \id)\kb{\Omega}{\Omega}).
\end{equation}

\begin{thm}[Stinespring's dilation theorem]
    Let $\Lambda:\cB(\cH)\to \cB(\cH)$ be a quantum operator. Then there exists a Hilbert space $\cH'$ and a unitary operator $U\in \cB(\cH\otimes \cH')$ such that $\forall \rho \in \cD(\cH)$,
    \begin{equation}
        \Lambda(\rho)=\Tr_{\cH'}(U(\rho\otimes \phi)U^\dagger)
    \end{equation}
    where $\phi$ is some fixed (pure) state in $\cH'$.
\end{thm}
That is, to perform a quantum operation we can couple to an ancilla, perform the unitary operation, and trace over the degrees of freedom in the ancilla $\cH'$.

Stinespring's dilation theorem is a result from operator theory, but we'll see shortly that there are two more equivalent and relevant formulations, known as the Kraus Representation Theorem and the C-J isomorphism. We'll discuss this first one today.

\begin{thm}[Kraus Rep'n Theorem]
    A linear map $\Lambda:\cM(\cH)\to \cB(\mathcal{K})$ is CP if and only if
    \begin{equation}\label{krausdecomp}
        \Lambda(\rho)=\sum_{k=1}^r A_k \rho A_k^\dagger
    \end{equation}
    where $\set{A_k}_{k=1}^r$ is a finite set of linear operators in $\cB(\cH,\mathcal{K}).$ Moreover it is TP if and only if
    \begin{equation}\label{krausunitaryish}
        \sum_{k=1}^r A_k^\dagger A_k = I_{\cH}.
    \end{equation}
\end{thm}
\begin{proof}
    We start by proving that the latter holds if the map is trace preserving and \ref{krausdecomp} holds. That is, trace preserving tells us that
    \begin{align*}
        1 &= \Tr \Lambda(\rho)\\
            &= \Tr \sum_k A_k \rho A_k^\dagger\\
            &= \sum_k \Tr(A_k \rho A_k^\dagger)\\
            &= \sum \Tr(A_k^\dagger A_k \rho)\\
            &= \Tr\paren{(\sum_k A_k^\dagger A_k)\rho} \,\forall \rho\\
            &\implies \sum_k A_k^\dagger A_k= I_{\cH}.
    \end{align*}
    Here, we have done nothing other than use definitions and the linearity and cyclic property of the trace.
\end{proof}
\subsection*{Kraus Rep'n Thm $\equiv$ restatement of Stin. D. Thm.}
We can think of the Kraus Representation theorem as a local implementation of the Stinespring Dilation Theorem. WLOG assume $\phi\equiv \kb{\phi}{\phi} \in \cD(\cH')$. Let $\set{\ket{e_k}}_k$ be an onb for $\cH'$. By Kraus, we can write
\begin{equation}
    \Lambda(\rho)=\sum_k \bra{e_k}U(\rho\otimes \phi) U^\dagger \ket{e_k}=\sum_k A_k \rho A_k^\dagger.
\end{equation}
with $\phi$ defined as above. That is, $\Lambda(\rho)=\Tr_{\cH'}(U(\rho \otimes \phi)U^\dagger)$. We define
\begin{equation}
    A_k:= \bra{e_k}U \ket{\phi}
\end{equation}
where $U\in \cB(\cH \otimes \cH')$ and it follows that
\begin{align*}
    \sum_k A_k^\dagger A_k &= \sum_k \bra{\phi}U^\dagger \kb{e_k}{e_k} U \ket{\phi}\\
    &= \braket{\phi}{\phi}I_{\cH} = I_{\cH}.
\end{align*}
We call the $A_k$ Kraus operators. Some of the details are an exercise to fill in later.

\subsection*{Choi-Jamilkowski (C-J) isomorphism} We saw that $\Lambda:\cB(\cH)\to \cB(\mathcal{K})$ where $\cH\simeq \CC^d, \mathcal{K}\simeq \CC^{d'}$ is CP iff $J(\Lambda)=(\Lambda \otimes \id_d)\kb{\Omega}{\Omega} \geq 0$. In fact, it turns out that $\exists$ an isomorphism between linear maps and positive operators. This is a great result, since positive operators are much nicer to work with.
\begin{thm}
    The following equation provides a bijection between linear maps $\Lambda:\cM_d \to \cM_{d'}$ and operators $J \in \cB(\CC^{d'}\otimes \CC^d)$, with $J$ defined as follows:
    \begin{equation}\label{cjforwards}
        J\equiv (\Lambda \otimes \id_d)\kb{\Omega}{\Omega}
    \end{equation}
    and
    \begin{equation}\label{cjbackwards}
        \Tr(A\Lambda(B))=d \Tr (J(A\otimes B^T))
    \end{equation}
    $\forall A \in \cM_{d'},B\in \cM_d$.
    The maps $\Lambda \to J \to \Lambda$ defined by \ref{cjforwards} and \ref{cjbackwards} are mutual inverses and lead to the following correspondence:
    \begin{enumerate}
        \item $\Lambda$ is CP $\iff J \geq 0$.
        \item $\Lambda$ is TP $\iff \Tr_A J = I_d/d$.
    \end{enumerate}
\end{thm}
\begin{proof}
    We'll first prove that \ref{cjforwards}$\to$\ref{cjbackwards}. The RHS of \ref{cjbackwards} is
    \begin{align*}
        \text{RHS}&=d\Tr (J(A\otimes B^T))\\
        &= d\Tr\paren{(\Lambda\otimes \id_d)\dyad{\Omega}(A\otimes B^T)}.
    \end{align*}
    Note we will need the concept of the \term{adjoint} $\Lambda^*$ of a map $\Lambda$ w.r.t. the Hilbert-Schmidt inner product. That is, if $\Lambda:\cB(\cH)\to \cB(\mathcal{K}),$ then $\Lambda^*:\cB(\mathcal{K})\to \cB(\cH)$ where
    \begin{equation}
        \Tr(A\Lambda(B)) = \Tr (\Lambda^*(A) B).
    \end{equation}
    Thus writing in terms of the adjoint, we have
    \begin{align*}
        \text{RHS}&=d\Tr((\Lambda\otimes \id_d)(\Omega)(A\otimes B^T))\\
        &= d\Tr((A\otimes B^T)(\Lambda\otimes \id_d) (\Omega))\\
        &= d\Tr((\Lambda^*(A)\otimes B^T)(\dyad{\Omega})).
        %&= d\Tr(\Omega(\Lambda^*\otimes \id_d)(A\otimes B^T))\\
        %&= d\Tr(\Omega(\Lambda^*(A)\otimes B^T))\\
        %&= d\Tr \paren{\Lambda^*(A)\otimes B^T (\kb{\Omega}{\Omega})}.
    \end{align*}
    Note this is slightly different from how it was presented in lecture. Here, I've used the cyclic property of the trace to switch the order of $J$ and $A\otimes B^T$, where I'm considering both as elements of $M_{d'}\otimes M_d$, and then I used the definition of the adjoint to change the $\Lambda$ into a $\Lambda^*$.%
        \footnote{It is also fairly clear that the adjoint of $\id$ is another identity operator on the appropriate space of matrices. Notice that $\Tr(A \id (B))=\Tr(AB)$ and $\Tr(A\id B)=\Tr(\id^*(A) B)$. For this to be true for all $A,B$, it must be that $\id^*=\id$, so the identity operator is self-adjoint. Thus we've sort of skipped a line here-- $(A\otimes B^T)(\Lambda\otimes \id_d)=\Lambda^*(A) \otimes \id^*(B^T)=\Lambda^*(A) \otimes B^T$. The result then follows.}
    Of course, we can split up the tensor product as
    \begin{align*}
        (\Lambda^*(A)\otimes B^T)\dyad{\Omega} &= (\Lambda^*(A) \otimes I)(I\otimes B^T)\dyad{\Omega}\\
        &= (\Lambda^*(A)\otimes I)(B\otimes I)\dyad{\Omega}\\
        &= (\Lambda^*(A)B \otimes I) \dyad{\Omega}\\
        &= (A \Lambda(B) \otimes I) \dyad{\Omega},
    \end{align*}
    where we have used the ricochet property in the second line to change a $B^T$ into a $B$ and turned $\Lambda^*$ back into a $\Lambda$. Finally, observe that this object (which after all is just $J(A\otimes B^T)$) lives in $M_{d'} \otimes M_d$. Let us denote a partial trace over the $M_{d'}$ subsystem by $\Tr_{d'}$ and over $M_d$, by $\Tr_d$. In this notation, we see that
    \begin{align*}
        d \Tr((A\Lambda(B)\otimes I) \dyad{\Omega}) &= \Tr\bkt{(A\Lambda(B)\otimes I)\sum_{i,j} \kb{i}{j} \otimes \kb{i}{j}}\\
            &= \Tr_{d'} \Tr_d\bkt{(A\Lambda(B)\otimes I)\sum_{i,j} \kb{i}{j} \otimes \kb{i}{j}}\\
            &= \Tr_{d'} \bkt{(A\Lambda(B)\otimes I)\sum_{i,j} \kb{i}{j} \delta_{ij}}\\
            &= \Tr_{d'} \bkt{\sum_i (A\Lambda(B))\dyad{i}}\\
            &= \Tr (A \Lambda(B)),
    \end{align*}
    where we recognize $\sum_i \dyad{i}$ as just the identity. We conclude that
    \begin{equation*}
        \Tr(A\Lambda(B))=d \Tr (J(A\otimes B^T)). \qedhere
    \end{equation*}
    % \begin{equation}
    %     d\Tr\bkt{(\Lambda^*(A)B\otimes I)\dyad{\Omega}}=\Tr(A\Lambda(B)),
    % \end{equation}
    % where we trace over the second system and use the adjoint to get back $\Lambda$.
\end{proof}
\begin{note}
In the statement of the CJ isomorphism, the notation is a little sneaky. In Eqn. \ref{cjbackwards}, the trace on the LHS is a trace over the space $\cM_{d'}$, since this is where $A$ and $\Lambda(B)$ live. But on the RHS, the trace is over $\cM_{d'}\otimes \cM_d$, since $J$ is a superoperator taking maps $\cM_{d'}\otimes \cM_d \to \cM_{d'}\otimes \cM_d$.

Note also that $J$ is \emph{not} a composition of linear maps. That is, it is not applying $\dyad{\Omega}$ and then some other map, because $\dyad{\Omega}$ is a map on operators in $\cM_d \otimes \cM_d$, whereas $J$ acts on operators in $\cM_{d'}\otimes \cM_d$.
\end{note}